---
title: "Yeilds & Fertilizer Use Efficiency"
author: "Tyler Cobian"
date: "3/11/2019"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE)
library(devtools)
library(gganimate)
library(devtools)
library(tidyverse)
library(RColorBrewer)
library(gifski)
library(png)



```

```{r}
fertilizer<- read_csv("fertilizer.csv")
grain<- read_csv("grain.csv") 

```
1. Patterns in yeild over time
In “grains.csv”, there is a column for yield: bushels/acre. As discussed in Grassini et al. 2013 (from week6), there are 6 possible statistical models to describe crop yield trends over time.

a. Create a finalized plot (including title and caption) of yield vs. time for each of the grains in the dataset.

```{r}
yeilds_graph<- ggplot(grain, aes(x = year, y = yield))+
  geom_line(aes(color = commodity), size = 0.6, alpha = 0.8)+
  facet_wrap(~commodity, scales = "free")+
  theme_classic()+
  labs(title = "Changes in Yeilds by Commodity", caption = "Yeild growth in the United States for corn, barley, oats, and sorghum
       from 1866 to 2018 measured in bushels per acre.", y = "Yield (bushels per acre)")+
  transition_reveal(id = yield, along = year)

animate(yeilds_graph, nframes = 24, renderer = gifski_renderer("Cobian_Tyler_animated.gif"))
  

yeilds_graph

```

2. Create a finalized plot (including title and caption) of fertilizer use vs time for each of the crops in the dataset.

```{r}
fert_graph<- ggplot(fertilizer, aes(x = year, y = ppa))+
  geom_point(aes(color = fertilizer), size = 0.5, alpha = 1)+
  facet_wrap(~crop)+
  scale_fill_brewer("Blues")+
  labs(title = "Fertilizer Application by Crop",
       caption = "The increase in nitrogen, phosphate, and potassium fertilizer In the United States 
       between 1964 and 2017 applied to corn, soybeans, and wheat in pounds per acre.",
       y = "Pounds per Acre")+
  theme(plot.title = element_text(hjust = 0.5, size = 10),
        plot.caption = element_text(hjust = 0.5, size = 8))
 




fert_graph

```
3. For corn, create a finalized plot (including title and caption) of yield vs fertilizer ppa, separated by the type of fertilizer. Make the color of the dots correspond to year.


```{r, include=FALSE}
fert_1<- fertilizer %>% 
  filter(crop == "Corn")
fert_1

grain_1<- grain %>% 
  filter(commodity == "Corn")
grain_1

yield_grain<- merge(fert_1, grain_1) %>% 
  select(year, crop, ppa, yield, fertilizer)

yield_grain
```


```{r}
# make a new df with fertilizer usage and yeild for corn
yg_graph<- ggplot(yield_grain)+
  geom_line(aes(x = yield, y = ppa, colour = year))+
  facet_wrap(~fertilizer)+
  labs(title = "Corn Yield Gains by Pound of Fertilizer Applied by Acre", caption = "Yeild gained from increased fertilizer application over time by fertilizer type.", y = "Yield (bushels per acre)", x = "Fertilizer (pounds per acre)")+
  theme(plot.caption = element_text(hjust = 0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent", colour = NA),
        plot.background = element_rect(fill = "transparent", colour = NA),
        legend.position = "none")
yg_graph

ggsave("yeild_return.png", yg_graph, bg = "transparent")

# make a CSV out of the yeild_grain

# write.csv(yield_grain, file = "returns_fertilizer.csv")



```

```{r, include=FALSE}
nitrogen_return<- yield_grain %>% 
  filter(fertilizer == "nitrogen")
nitrogen_return
write.csv(nitrogen_return, file = "nitrogen_return.csv")

phosphate_return<- yield_grain %>% 
  filter(fertilizer == "phosphate")
nitrogen_return
write.csv(nitrogen_return, file = "phosphate_return.csv")

potassium_return<- yield_grain %>% 
  filter(fertilizer == "potassium")
nitrogen_return
 write.csv(nitrogen_return, file = "potassium_return.csv")





```














